---
title: "PopGen"
author: "Amanda Charbonneau"
date: '`r Sys.Date()`'
output:
  html_document:
    code_folding: hide
    collapsed: no
    df_print: paged
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
  html_notebook:
    toc: yes
    toc_depth: 5
runtime: shiny
---


```{r, echo=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Install function for packages    
packages<-function(x){
  x<-as.character(match.call()[[2]])
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x,repos="http://cran.r-project.org")
    require(x,character.only=TRUE)
  }
}
packages("phangorn")
packages("adegenet")
packages("hierfstat")
packages("pegas")
packages("pheatmap")
packages("RColorBrewer")
packages("poppr")
packages("plyr")
packages("dplyr")
packages("ggplot2")
packages("shiny")


```

This script uses the input for STRUCTURE to do some population genetics.

```{r, echo=FALSE, message=FALSE}

#has population names and group names

ALLTHEALLELES <- read.structure(file = "../MungedData/1randSS.str", n.ind = 338, n.loc = 21, onerowperind = T, col.lab = 1, col.pop = 2, row.marknames = 1, NA.char = "-9", col.others = c(2,3))

strata(ALLTHEALLELES) <- data.frame(other(ALLTHEALLELES))

colnames(strata(ALLTHEALLELES)) <- c("Pop", "Group")

poppr_ATA <- as.genclone(ALLTHEALLELES)


groupcov <- read.csv("../Figures/GroupCoV.csv")
expcov <- read.csv("../Figures/ExpCoV.csv")
popcov <- read.csv("../Figures/PopCoV.csv")

```



```{r, echo=FALSE,message=FALSE}
#Genind has only individual information, genpop has population level info as well

poplevel <- genind2genpop(ALLTHEALLELES)

poplevelfreq <- tab(poplevel, freq=T)

write.csv(poplevelfreq, "../Figures/PopulationLevelAlleleFreq.csv")

grouplevel <- genind2genpop(x = ALLTHEALLELES, pop = ALLTHEALLELES@other$X[,2])
grouplevelfreq <- tab(grouplevel, freq=T)

write.csv(grouplevelfreq, "../Figures/GroupLevelAlleleFreq.csv")
```
#Summary of the marker data

##Individuals per Group

```{r, echo=FALSE}

table(strata(ALLTHEALLELES, ~Group))

```


##Data missingness.

There's a lot of missingness. At one point I ran a manual jackknife of STRUCTURE by both markers and populations to see if dropping any one made a difference. It didn't. I don't understand how it can be robust with that much missing, but it seems to be.

```{r, echo=FALSE, message=FALSE, fig.height=6}

bob <- info_table(ALLTHEALLELES, plot=TRUE, scaled=T)

#data_sets <- c(ALLTHEALLELES, FilterATA)

#inputPanel(
#  selectInput( "LocusSet", "Loci:", choices=c("All Loci"=1, "Filtered Loci"=2), selected=2)
#)

#renderPlot( info_table( data_sets[[as.numeric(input$LocusSet)]], plot = TRUE, scaled = T))
```



##Comparing expected and observed heterozygosity per marker
```{r, echo=FALSE, message=FALSE}

summary(ALLTHEALLELES)

ALLSUM <- adegenet::summary(ALLTHEALLELES)

stats::bartlett.test( list( ALLSUM$Hexp, ALLSUM$Hobs ) )

stats::t.test(ALLSUM$Hexp,ALLSUM$Hobs,pair=T,var.equal=TRUE,alter="greater")

```



## Summary stats per population

```{r, echo=FALSE, message=FALSE}

inputPanel(
  selectInput("popname", "Population:", choices = c("ADOL"=1,"AFFR"=2,"AROL"=3,"AUFI"=4,"BINY"=5,"CBBG"=6,"CBES"=7,"COAU"=8,"COOL"=9,"DAJO"=10,"DEES"=11,"ESNK"=12,"GHIL"=13,"GMIL"=14,"HCES"=15,"HMES"=16,"HZIL"=17,"IMES"=18,"MABG"=19,"MAES"=20,"MYJO"=21,"NCDE"=22,"NELO"=23,"PBFR"=24,"RABG"=25,"RACA"=26,"REIL"=27,"SAES"=28,"SPEU"=29,"TOBG"=30,"TYIL"=31,"WEAU"=32,"WMBG"=33,"ZYIL"=34)))

renderPrint({ adegenet::summary(seppop(ALLTHEALLELES)[[ as.numeric(input$popname) ]]) })

```


##Summary stats for all markers
```{r, echo=FALSE}
basic.stats(ALLTHEALLELES)

```

##Do markers appear to be in HWE across dataset?

```{r, echo=FALSE}

hw.test(ALLTHEALLELES, res='matrix', B=1000)

```
##Do markers appear to be in HWE by population?

Blue is population in departure from HWE at marker at a p < .05

```{r, echo=FALSE, message=FALSE, warning=FALSE}

#https://grunwaldlab.github.io/Population_Genetics_in_R/Locus_Stats.html

nanhwe.pop <- seppop(ALLTHEALLELES) %>% lapply(hw.test, B = 0)

nanhwe.mat <- sapply(nanhwe.pop, "[", i = TRUE, j = 3) # Take the third column with all rows

alpha  <- 0.05
newmat <- nanhwe.mat
newmat[newmat > alpha] <- 1

pheatmap(newmat)
```

##Comparing expected heterozygosity per population and per group

```{r, echo=FALSE}
Hs(poplevel)

Hs(grouplevel)

hspop <- as.data.frame(as.matrix(Hs(poplevel)))
colnames(hspop) <- "ExpHet"
hspop$Pop <- rownames(hspop)

hspop$Pop <- revalue(hspop$Pop, c(
                               "SPEU" = "SPNK",
                               "NELO" = "NEJS",
                               "RACA" = "RAJS"
                               ))




hsgroup <- as.data.frame(as.matrix(Hs(grouplevel)))
colnames(hsgroup) <- "ExpHet"
hsgroup$locals <- rownames(hsgroup)

hsgroup$locals <- revalue(hsgroup$locals, c(
                               "European" = "european",
                               "Oilseed"   = "oleifera",
                               "Rattail"  = "caudatus",
                               "Daikon" = "daikon",
                               "rostratus" = "pugioniformis"
                               ))


```

##Expected Heterozygosity vs Coeffecient of Variation

The weeds as a group have the highest heterozygosity, but a really low Coeffecient of variation. I wondered if they were anti-correlated.

It seems not.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
popHS_Cov <- full_join(hspop, popcov)
popHS_Cov$locals[popHS_Cov$Pop=="NCDE"] <- "raphNN"

ggplot(popHS_Cov, aes(ExpHet, COV, col=locals)) + geom_point() + geom_smooth(method = 'lm', se = F) + ggtitle("By Population")

write.csv(popHS_Cov, "../Figures/PopHS_Cov.csv")
groupHS_Cov <- full_join(hsgroup, groupcov)

ggplot(groupHS_Cov, aes(ExpHet, COV, col=locals)) + geom_point() + geom_smooth(method = 'lm', se = F) + ggtitle("By Group")
write.csv(groupHS_Cov, "../Figures/GroupHS_Cov.csv")

```


##PCA

A fun little pca that tries to make boundries around pops (using the dudi.pca function from ade4). Labels seem unmoveable though, so I can't get it to format nicely for use. Looks essentially the same as the paper one, even though it's using the dataset in a completely different format/program/etc. So that's nice.

```{r, echo=FALSE}
testpca <- scaleGen(ALLTHEALLELES, NA.method="mean")
dim(testpca)

pca1 <- dudi.pca(testpca,cent=FALSE,scale=FALSE,scannf=FALSE,nf=3)
#barplot(pca1$eig[1:50],main="PCA eigenvalues", col=heat.colors(50))


col <- funky(34)
s.class(pca1$li, pop(ALLTHEALLELES),xax=1,yax=2, col=transp(col,.6), axesell=FALSE,
        cstar=1, cpoint=1, grid=FALSE, clabel = .5)


#colorplot(pca1$li, pca1$li, transp=TRUE, cex=3, xlab="PC 1", ylab="PC 2")
#title("PCA of microbov dataset\naxes 1-2")
#abline(v=0,h=0,col="grey", lty=2)
```


```{r, echo=FALSE, eval=FALSE}

#Playing with adegnet tutorial: An introduction to adegenet 2.1.0 6/14/2017

ca1 <- dudi.coa(tab(poplevel),scannf=FALSE,nf=3)
barplot(ca1$eig,main="Correspondance Analysis eigenvalues",
        col=heat.colors(length(ca1$eig)))

 s.label(ca1$li, sub="CA 1-2",csub=2)


set.seed(1)
s.label(ca1$li*1.2, sub="CA 1-2",csub=2, clab=0, cpoint="")
textplot(ca1$li[,1], ca1$li[,2], words=popNames(poplevel),
         cex=1.4, new=FALSE, xpd=TRUE)
add.scatter.eig(ca1$eig,nf=3,xax=1,yax=2,posi="bottomright")
```

##Populations clustered by genetic distance. Euclidean method.

Makes some expected groupings: landra, weeds(BINY, AUFI, NCDE), most of the crops
Has Australian weeds clustered with native RRR's instead of other weeds. Sort of like STRUCTURE

```{r, echo=FALSE}
sampleDists <- dist.genpop(poplevel, method = 1, diag = T, upper = T)
sampleDistMatrix <- as.matrix(sampleDists)
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows="euclidean",
         clustering_distance_cols="euclidean",
         cluster_rows = T,
         cluster_cols = T,
         col=colors)
```

##Same clustering method, just using groups instead of pops.


```{r, echo=FALSE}
sampleDists <- dist.genpop(grouplevel, method = 1, diag = T, upper = T)
sampleDistMatrix <- as.matrix(sampleDists)
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows="euclidean",
         clustering_distance_cols="euclidean",
         cluster_rows = T,
         cluster_cols = T,
         col=colors)
```

##Fstatistics

These are the three Fstatistics for the overall dataset. You are supposed to read each value as 'column/row'. These are calculated as Nei's pairwise Fst between all pairs of populations using a genind object. Heretozygosities are weighted by group sizes. I have zero idea what it is doing with missing data points, or if these are remotely accurate.

```{r, echo=FALSE}
fstat(ALLTHEALLELES)
```

Pairwise Fst for all populations. This is output as a csv file, which is much easier to read.

This is the Fst numbers, just as a clustered heatmap so you can see patterns more easily.

```{r, echo=FALSE, warning=FALSE}

popfst <- pairwise.fst(ALLTHEALLELES)

attr(popfst, "Labels") <- unique(as.character(ALLTHEALLELES$pop))

write.csv(as.data.frame(as.matrix(popfst)), "../Figures/PairwiseFst.csv", sep = "'", quote = F, row.names = TRUE, col.names = TRUE )

```



```{r, echo=FALSE}

pheatmap(as.matrix(popfst), cluster_rows = T, cluster_cols = T,
         clustering_distance_rows = "euclidean",
         clustering_distance_cols = "euclidean" )
```



##Allele by Pop

I don't remember why I made this, but it's a clustered heatmap of the population level allele frequencies. The columns are alleles. A lot of alleles are at low frequency, and then a handful at very high frequency. Columns where it's all blue with one red or yellow row indicate private alleles. Mostly red/yellow columns indicate alleles that are common. Columnes that are mostly blue indicate low frequency alleles.


```{r, echo=FALSE}
pheatmap(poplevelfreq, cluster_rows=TRUE, show_rownames=TRUE, cluster_cols=T, show_colnames = F)
```
##Allele by Group

Again, same type of plot, just using groups instead of populations. There are a few alleles private to groups, and a lot of very low frequency alleles.

```{r, echo=FALSE}
pheatmap(grouplevelfreq, cluster_rows=TRUE, show_rownames=TRUE, cluster_cols=T, show_colnames = F)
```


```{r, echo=FALSE, eval=FALSE}

## K means clustering

#Just to see what happens. I have no idea what this figure means.

disana <- dapc(ALLTHEALLELES, pop=ALLTHEALLELES@other$X[,2], n.pca=40, n.da=15, scale=FALSE,
          truenames=TRUE, var.contrib=TRUE, var.loadings=FALSE, pca.info=TRUE,
          pca.select=c("nbEig","percVar"), perc.pca=NULL)


scatter(disana, cell=2, pch="")


kmeans <- find.clusters(ALLTHEALLELES, n.pca=40, n.clust=NULL,
              stat="BIC", choose.n.clust=FALSE,
              criterion="diffNgroup",
              max.n.clust=20, n.iter=1e5,
              n.start=10, scale=FALSE, truenames=TRUE)


```


```{r, message=T, error=TRUE, warning=TRUE, fig.height=8}



inputPanel(
  selectInput("Grouplevel", "Grouping:", choices = c("Pop"="~Pop", "Group"="~Group", "Group/Pop"="~Group/Pop" ), selected = "~Group"),
  selectInput("miss", "Missing Data options:", choices=c("loci", "ignore", "zero", "mean", "geno"), selected = "geno"),
  sliderInput("cutoff", "Missingness Cutoff:", min=0, max=1, value=.05, step=.05),
  checkboxGroupInput("goodloci", "Choose Loci to use:", choices = c("DWRD_124", "DWRD_112", "DWRD_61",  "DWRD_177", "DWRD_107",
"DWRD_123", "DWRD_121", "DWRD_158", "DWRD_48",  "DWRD_180",
"DWRD_97",  "DWRD_205", "DWRD_27",  "Bn26a",    "BRMS005", "Na10H06",
"Ra1H08", "Ra2E11", "Na14E08", "Bn35d", "Na12E05"), selected = c("DWRD_124", "DWRD_112", "DWRD_61",  "DWRD_177", "DWRD_107",
"DWRD_123", "DWRD_121", "DWRD_158", "DWRD_48",  "DWRD_180",
"DWRD_97",  "DWRD_205", "DWRD_27",  "Bn26a",    "BRMS005", "Na10H06",
"Ra1H08", "Ra2E11", "Na14E08", "Bn35d", "Na12E05")),
checkboxGroupInput("goodgroups", "Choose groups to use:", choices=c("RRR_Native_West", "RRR_Native_East", "RRR_Weed", "Rr_Landra", "RS_Daikon", "RS_European", "RS_Oilseed", "RS_Rattail", "Rostratus" ), selected = c("RRR_Native_West", "RRR_Native_East", "RRR_Weed", "Rr_Landra", "RS_Daikon", "RS_European", "RS_Oilseed", "RS_Rattail", "Rostratus")))

renderText( { goodgroups <- c(input$goodgroups)
              goodloci <- c(input$goodloci)
              FilterATA <- ALLTHEALLELES[loc=goodloci]
              Allpops_filter_ATA <- as.genclone(FilterATA)
              poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = c(input$goodgroups))
  capture.output(
    poppr.amova(poppr_filter_ATA, as.formula(input$Grouplevel), filter = T,
                missing = input$miss, cutoff= as.numeric(input$cutoff)),
                type=c("message"))})


renderPlot({  RRR_Native_West <- c("AFFR", "MAES","HCES", "HMES","IMES", "DEES")
RRR_Native_East <- c("GHIL", "HZIL", "REIL", "TYIL", "ZYIL")
RRR_Weed <- c("AUFI", "BINY", "COAU", "WEAU", "NCDE")
Rr_Landra <- c("CBES", "SAES", "PBFR")
RS_Daikon <- c("MYJO", "NELO", "TOBG", "WMBG")
RS_European <- c("DAJO", "ESNK", "SPEU", "CBBG")
RS_Oilseed <- c("AROL", "COOL", "ADOL")
RS_Rattail <- c("RABG", "RACA", "MABG")
Rostratus <- c("GMIL")
goodloci <- c(input$goodloci)
              goodgroups <- c(input$goodgroups)
              FilterATA <- ALLTHEALLELES[loc=goodloci]
              Allpops_filter_ATA <- as.genclone(FilterATA)
              poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = goodgroups)
              AMOVA_ATA <- poppr.amova(poppr_filter_ATA, as.formula(input$Grouplevel), filter = T,
                                       missing = input$miss, cutoff= as.numeric(input$cutoff))
              amova.test <- randtest( AMOVA_ATA ) # Test for significance
              plot(amova.test) })

renderPrint( {goodloci <- c(input$goodloci)
              FilterATA <- ALLTHEALLELES[loc=goodloci]
              Allpops_filter_ATA <- as.genclone(FilterATA)
              poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = c(input$goodgroups))
              AMOVA_ATA <- poppr.amova(poppr_filter_ATA, as.formula(input$Grouplevel), filter = T, missing = input$miss, cutoff= as.numeric(input$cutoff))
  randtest( AMOVA_ATA )})

renderPrint( {goodloci <- c(input$goodloci)
              FilterATA <- ALLTHEALLELES[loc=goodloci]
              Allpops_filter_ATA <- as.genclone(FilterATA)
              poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = c(input$goodgroups))
              poppr.amova(poppr_filter_ATA, as.formula(input$Grouplevel), filter = T, missing = input$miss, cutoff= as.numeric(input$cutoff)) })

```
