---
title: "PopGen"
author: "Amanda Charbonneau"
date: '`r Sys.Date()`'
output:
  html_document:
    code_folding: hide
    collapsed: no
    df_print: paged
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
  html_notebook:
    toc: yes
    toc_depth: 5
---


```{r, echo=FALSE, message=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.path="../Figures/", fig.keep="last")

# Install function for packages    
packages<-function(x){
  x<-as.character(match.call()[[2]])
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x,repos="http://cran.r-project.org")
    require(x,character.only=TRUE)
  }
}
packages("phangorn")
packages("adegenet")
packages("hierfstat")
packages("pegas")
packages("pheatmap")
packages("RColorBrewer")
packages("poppr")
packages("plyr")
packages("dplyr")
packages("ggplot2")
packages("shiny")
packages("StatMatch")

amovaReps <- 500

```

```{r}
#Set up plotting colors

## Add an alpha value to a colour
add.alpha <- function(col, alpha=.7){
  if(missing(col))
    stop("Please provide a vector of colours.")
  apply(sapply(col, col2rgb)/255, 2,
        function(x)
          rgb(x[1], x[2], x[3], alpha=alpha))  
}

col_pal <- brewer.pal(9, "Set1")
col6 <- brewer.pal(6, "Set2")
col_pal <- c(col_pal[c(1:5)], col6[6], col_pal[c(8,9)], "gray20", "black")

ALLTHECOLORS <- add.alpha(col_pal)

NonNatives <- ALLTHECOLORS[5]
Daikon <- ALLTHECOLORS[6]
European <- ALLTHECOLORS[1]
lanmar <- ALLTHECOLORS[7]
Oilseed <- ALLTHECOLORS[3]
raphNatE <- ALLTHECOLORS[2]
raphNatW <- ALLTHECOLORS[4]
Rattail <- ALLTHECOLORS[8]
rostratus <-  ALLTHECOLORS[10]
```


This script uses the input for STRUCTURE to do some population genetics.

```{r, echo=FALSE, message=FALSE}

#has population names and group names

ALLTHEALLELES <- read.structure(file = "../MungedData/1randSS.str", n.ind = 338, n.loc = 21, onerowperind = T, col.lab = 1, col.pop = 2, row.marknames = 1, NA.char = "-9", col.others = c(2,3))

adegenet::strata(ALLTHEALLELES) <- read.csv("../OriginalData/PopStructureAmova.csv", header = FALSE) 

adegenet::splitStrata(ALLTHEALLELES) <- ~Species/Group/Pop

groupcov <- read.csv("../Figures/GroupCoV.csv")
expcov <- read.csv("../Figures/ExpCoV.csv")
popcov <- read.csv("../Figures/PopCoV.csv")

#goodloci <- c( "DWRD_124", "DWRD_112", "DWRD_61", "DWRD_177", "DWRD_107",
#"DWRD_123", "DWRD_121", "DWRD_158", "DWRD_48", "DWRD_180",
#"DWRD_97", "DWRD_205", "DWRD_27", "Bn26a", "BRMS005",
#"Ra1H08", "Ra2E11", "Bn35d", "Na12E05" )
```


```{r, echo=FALSE,message=FALSE}
#Genind has only individual information, genpop has population level info as well

poplevel <- genind2genpop(ALLTHEALLELES)

poplevelfreq <- tab(poplevel, freq=T)

write.csv(poplevelfreq, "../Figures/PopulationLevelAlleleFreq.csv")

grouplevel <- genind2genpop(x = ALLTHEALLELES, pop = ALLTHEALLELES@other$X[,2])
grouplevelfreq <- tab(grouplevel, freq=T)

write.csv(grouplevelfreq, "../Figures/GroupLevelAlleleFreq.csv")
```
# Summary of the marker data

## Individuals per Group

```{r, echo=FALSE}

table(adegenet::strata(ALLTHEALLELES, ~Group))

```


## Data missingness.

There's a lot of missingness. At one point I ran a manual jackknife of STRUCTURE by both markers and populations to see if dropping any one made a difference. It didn't. I don't understand how it can be robust with that much missing, but it seems to be.

```{r missingdata, echo=FALSE, message=FALSE, fig.height=6}

bob <- info_table(ALLTHEALLELES, plot=TRUE, scaled=T)

#data_sets <- c(ALLTHEALLELES, FilterATA)

#inputPanel(
#  selectInput( "LocusSet", "Loci:", choices=c("All Loci"=1, "Filtered Loci"=2), selected=2)
#)

#renderPlot( info_table( data_sets[[as.numeric(input$LocusSet)]], plot = TRUE, scaled = T))
```



## Comparing expected and observed heterozygosity per marker
```{r, echo=FALSE, message=FALSE}

summary(ALLTHEALLELES)

ALLSUM <- adegenet::summary(ALLTHEALLELES)

stats::bartlett.test( list( ALLSUM$Hexp, ALLSUM$Hobs ) )

stats::t.test(ALLSUM$Hexp,ALLSUM$Hobs,pair=T,var.equal=TRUE,alter="greater")

```



## Summary stats per population

```{r, echo=FALSE, message=FALSE, eval=FALSE}

inputPanel(
  selectInput("popname", "Population:", choices = c("ADOL"=1,"AFFR"=2,"AROL"=3,"AUFI"=4,"BINY"=5,"CBBG"=6,"CBES"=7,"COAU"=8,"COOL"=9,"DAJO"=10,"DEES"=11,"ESNK"=12,"GHIL"=13,"GMIL"=14,"HCES"=15,"HMES"=16,"HZIL"=17,"IMES"=18,"MABG"=19,"MAES"=20,"MYJO"=21,"NCDE"=22,"NELO"=23,"PBFR"=24,"RABG"=25,"RACA"=26,"REIL"=27,"SAES"=28,"SPEU"=29,"TOBG"=30,"TYIL"=31,"WEAU"=32,"WMBG"=33,"ZYIL"=34)))

renderPrint({ adegenet::summary(seppop(ALLTHEALLELES)[[ as.numeric(input$popname) ]]) })

```


## Summary stats for all markers
```{r, echo=FALSE}
basic.stats(ALLTHEALLELES)

```

## Do markers appear to be in HWE across dataset?

```{r, echo=FALSE}

hw.test(ALLTHEALLELES, res='matrix', B=1000)

```
## Do markers appear to be in HWE by population?

Blue is population in departure from HWE at marker at a p < .05

```{r HWE_pop_marker, echo=FALSE, message=FALSE, warning=FALSE}

#https://grunwaldlab.github.io/Population_Genetics_in_R/Locus_Stats.html

nanhwe.pop <- seppop(ALLTHEALLELES) %>% lapply(hw.test, B = 0)

nanhwe.mat <- sapply(nanhwe.pop, "[", i = TRUE, j = 3) # Take the third column with all rows

alpha  <- 0.05
newmat <- nanhwe.mat
newmat[newmat > alpha] <- 1

pheatmap(newmat)
```

## Comparing expected heterozygosity per population and per group

```{r, echo=FALSE}
Hs(poplevel)

Hs(grouplevel)

hspop <- as.data.frame(as.matrix(Hs(poplevel)))
colnames(hspop) <- "ExpHet"
hspop$Pop <- rownames(hspop)

hspop$Pop <- plyr::revalue(hspop$Pop, c(
                               "SPEU" = "SPNK",
                               "NELO" = "NEJS",
                               "RACA" = "RAJS"
                               ))




hsgroup <- as.data.frame(as.matrix(Hs(grouplevel)))
colnames(hsgroup) <- "ExpHet"
hsgroup$locals <- rownames(hsgroup)

hsgroup$locals <- revalue(hsgroup$locals, c(
                               "European" = "european",
                               "Oilseed"   = "oleifera",
                               "Rattail"  = "caudatus",
                               "Daikon" = "daikon",
                               "rostratus" = "pugioniformis"
                               ))


```

##Expected Heterozygosity vs Coeffecient of Variation

The weeds as a group have the highest heterozygosity, but a really low Coeffecient of variation. I wondered if they were anti-correlated.

It seems not.

```{r hetero_CoV, echo=FALSE, message=FALSE, warning=FALSE}
popHS_Cov <- full_join(hspop, popcov)
popHS_Cov$locals[popHS_Cov$Pop=="NCDE"] <- "raphNN"

ggplot(popHS_Cov, aes(ExpHet, COV, col=locals)) + geom_point() + geom_smooth(method = 'lm', se = F) + ggtitle("By Population")

write.csv(popHS_Cov, "../Figures/PopHS_Cov.csv")
groupHS_Cov <- full_join(hsgroup, groupcov)

ggplot(groupHS_Cov, aes(ExpHet, COV, col=locals)) + geom_point() + geom_smooth(method = 'lm', se = F) + ggtitle("By Group")
write.csv(groupHS_Cov, "../Figures/GroupHS_Cov.csv")

```


## PCA

A fun little pca that tries to make boundries around pops (using the dudi.pca function from ade4). Labels seem unmoveable though, so I can't get it to format nicely for use. Looks essentially the same as the paper one, even though it's using the dataset in a completely different format/program/etc. So that's nice.

```{r dudiPCA, echo=FALSE}
testpca <- scaleGen(ALLTHEALLELES, NA.method="mean")
dim(testpca)

pca1 <- dudi.pca(testpca,cent=FALSE,scale=FALSE,scannf=FALSE,nf=3)
#barplot(pca1$eig[1:50],main="PCA eigenvalues", col=heat.colors(50))


col <- funky(34)
s.class(pca1$li, pop(ALLTHEALLELES),xax=1,yax=2, col=transp(col,.6), axesell=FALSE,
        cstar=1, cpoint=1, grid=FALSE, clabel = .5)


#colorplot(pca1$li, pca1$li, transp=TRUE, cex=3, xlab="PC 1", ylab="PC 2")
#title("PCA of microbov dataset\naxes 1-2")
#abline(v=0,h=0,col="grey", lty=2)
```


```{r, echo=FALSE, eval=FALSE}

#Playing with adegnet tutorial: An introduction to adegenet 2.1.0 6/14/2017

ca1 <- dudi.coa(tab(poplevel),scannf=FALSE,nf=3)
barplot(ca1$eig,main="Correspondance Analysis eigenvalues",
        col=heat.colors(length(ca1$eig)))

 s.label(ca1$li, sub="CA 1-2",csub=2)


set.seed(1)
s.label(ca1$li*1.2, sub="CA 1-2",csub=2, clab=0, cpoint="")
textplot(ca1$li[,1], ca1$li[,2], words=popNames(poplevel),
         cex=1.4, new=FALSE, xpd=TRUE)
add.scatter.eig(ca1$eig,nf=3,xax=1,yax=2,posi="bottomright")
```

## Populations clustered by genetic distance. Euclidean method.

Makes some expected groupings: landra, weeds(BINY, AUFI, NCDE), most of the crops
Has Australian weeds clustered with native RRR's instead of other weeds. Sort of like STRUCTURE

```{r GenDist_Pop, echo=FALSE}
sampleDists <- dist.genpop(poplevel, method = 1, diag = T, upper = T)
sampleDistMatrix <- as.matrix(sampleDists)
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows="euclidean",
         clustering_distance_cols="euclidean",
         cluster_rows = T,
         cluster_cols = T,
         col=colors)
```

## Same clustering method, just using groups instead of pops.


```{r GenDist_Group, echo=FALSE}
sampleDists <- dist.genpop(grouplevel, method = 1, diag = T, upper = T)
sampleDistMatrix <- as.matrix(sampleDists)
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows="euclidean",
         clustering_distance_cols="euclidean",
         cluster_rows = T,
         cluster_cols = T,
         col=colors)
```

## Fstatistics

These are the three Fstatistics for the overall dataset. You are supposed to read each value as 'column/row'. These are calculated as Nei's pairwise Fst between all pairs of populations using a genind object. Heretozygosities are weighted by group sizes. I have zero idea what it is doing with missing data points, or if these are remotely accurate.

```{r, echo=FALSE}
fstat(ALLTHEALLELES)
```

Pairwise Fst for all populations. This is output as a csv file, which is much easier to read.

This is the Fst numbers, just as a clustered heatmap so you can see patterns more easily.

```{r, echo=FALSE, warning=FALSE}

popfst <- pairwise.fst(ALLTHEALLELES)#[ loc=goodloci ])

attr(popfst, "Labels") <- unique(as.character(ALLTHEALLELES$pop))

write.csv(as.data.frame(as.matrix(popfst)), "../Figures/PairwiseFst.csv", sep = "'", quote = F, row.names = TRUE, col.names = TRUE )

```



```{r FstPop, echo=FALSE}

pheatmap(as.matrix(popfst), cluster_rows = T, cluster_cols = T,
         clustering_distance_rows = "euclidean",
         clustering_distance_cols = "euclidean" )
```

```{r FstPopGroup, echo=FALSE}
grouppopfst <- as.data.frame(as.matrix(popfst))

grouppopfst$group <- unique(as.character(ALLTHEALLELES$pop))
#grouppopfst$pop <- unique(as.character(ALLTHEALLELES$pop))

grouppopfst$group <- as.factor(plyr::revalue(grouppopfst$group, c("COAU"="WeedRrr", "WEAU"="WeedRrr", 
"AUFI"="WeedRrr", "BINY"="WeedRrr",
"SAES"="Rrlandra", "MAES"="NativeRrrW",
"PBFR"="Rrlandra", "ESNK"="European",
"AFFR"="NativeRrrW", "AROL"="Oilseed",
"RACA"="Rattail", "COOL"="Oilseed",
"ADOL"="Oilseed", "NELO"="Daikon",
"SPEU"="European", "GMIL"="Rostratus",
"TYIL"="NativeRrrE", "REIL"="NativeRrrE",
"CBES"="Rrlandra", "MABG"="Rattail",
"TOBG"="Daikon", "CBBG"="European",
"WMBG"="Daikon", "RABG"="Rattail",
"DAJO"="European", "MYJO"="Daikon",
"NCDE"="WeedRrr", "GHIL"="NativeRrrE",
"HZIL"="NativeRrrE", "ZYIL"="NativeRrrE",
"DEES"="NativeRrrW", "HCES"="NativeRrrW",
"HMES"="NativeRrrW", "IMES"="NativeRrrW")))

groupnames <- data.frame(grouppopfst[,35])
rownames(groupnames) <- unique(as.character(ALLTHEALLELES$pop))
colnames(groupnames) <- "group"
  
groupcolors <- list( group = c(Daikon=paste(Daikon), European=paste(European), 
          NativeRrrE=paste(raphNatE), NativeRrrW=paste(raphNatW), 
          Oilseed=paste(Oilseed), Rattail=paste(Rattail), Rostratus=paste(rostratus),
          Rrlandra=paste(lanmar), WeedRrr=paste(NonNatives)))

pheatmap( as.matrix( grouppopfst[,1:34] ), annotation_col = groupnames, 
          annotation_row = groupnames , cluster_cols = T, cluster_rows = T,
          annotation_colors = groupcolors)

```

## Allele by Pop

I don't remember why I made this, but it's a clustered heatmap of the population level allele frequencies. The columns are alleles. A lot of alleles are at low frequency, and then a handful at very high frequency. Columns where it's all blue with one red or yellow row indicate private alleles. Mostly red/yellow columns indicate alleles that are common. Columns that are mostly blue indicate low frequency alleles.


```{r AllelePop, echo=FALSE}
pheatmap(poplevelfreq, cluster_rows=TRUE, show_rownames=TRUE, cluster_cols=T, show_colnames = F)
```
##Allele by Group

Again, same type of plot, just using groups instead of populations. There are a few alleles private to groups, and a lot of very low frequency alleles.

```{r AlleleGroup, echo=FALSE}
pheatmap(grouplevelfreq, cluster_rows=TRUE, show_rownames=TRUE, cluster_cols=T, show_colnames = F)
```


```{r, echo=FALSE, eval=FALSE}

## K means clustering

#Just to see what happens. I have no idea what this figure means.

disana <- dapc(ALLTHEALLELES, pop=ALLTHEALLELES@other$X[,2], n.pca=40, n.da=15, scale=FALSE,
          truenames=TRUE, var.contrib=TRUE, var.loadings=FALSE, pca.info=TRUE,
          pca.select=c("nbEig","percVar"), perc.pca=NULL)


scatter(disana, cell=2, pch="")


kmeans <- find.clusters(ALLTHEALLELES, n.pca=40, n.clust=NULL,
              stat="BIC", choose.n.clust=FALSE,
              criterion="diffNgroup",
              max.n.clust=20, n.iter=1e5,
              n.start=10, scale=FALSE, truenames=TRUE)


```

# AMOVA

### Overall group AMOVA

#### Ignore missing data

For the basic AMOVA, I tried both dropping the two worst microsats entirely, and leaving them in. There is basically no difference between the results, so I am ignoring all missingness. (To remove 2 bad markers, uncomment "[ loc=goodloci ]" throughout the script, and uncomment the list of "goodloci" in the thrid code chunk). 

I am looking for difference ~Groups/Pop. Groups are: landra, western raphanistrum, eastern raphanistrum, weeds, European, Daikon, Rattail, Oilseed, and rostratus.

In the results "sample" always refers to the smallest unit of the population hierarchy, which is population here.

```{r AmovaAll, message=T, error=TRUE, warning=TRUE, fig.height=8}

#Defaults to ADE4 AMOVA implementation

FilterATA <- ALLTHEALLELES#[ loc=goodloci ]

Allpops_filter_ATA <- as.genclone( FilterATA )

Allpopsamova <- poppr.amova( Allpops_filter_ATA, ~Group/Pop, filter = T, 
                             missing = "ignore", within = FALSE )

Allpopsamova

amova.test <- randtest( Allpopsamova, nrepet = amovaReps) # Test for significance
plot(amova.test)
amova.test

## The genetic differentiation components measured in AMOVA are really covariances among alleles, not variances, and negative values should be interpreted in that context.  That said, such strongly negative values for between-population differentiation are problematic.  However, obligately outcrossing organisms, or those that suffer from severely reduced viability due to inbreeding, can have substantially negative values for the within-population inbreeding coefficient.  If your analysis isn't estimating the inbreeding coefficient separately, or is estimating it incorrectly (e.g. inflated estimates of inbreeding due to the presence of null alleles), I wonder if these effects could be masquerading as negative allelic covariances between populations in the absence of population structure.

```

This is the distribution of N for groups that AMOVA was using. It's not super balanced, but it's not the worst.

```{r}

table(adegenet::strata(ALLTHEALLELES, ~Group))
#filterstrata <- adegenet::strata(Allpops_filter_ATA) %>%     
#  group_by(Species, Group) %>%
#  summarize(Count = n())
#
#filterstrata 
```

### Hypothesis testing AMOVAs

```{r}
RRR_Native_West <- c("AFFR", "MAES","HCES", "HMES","IMES", "DEES")
RRR_Native_East <- c("GHIL", "HZIL", "REIL", "TYIL", "ZYIL")
RRR_Weed <- c("AUFI", "BINY", "COAU", "WEAU", "NCDE")
Rr_Landra <- c("CBES", "SAES", "PBFR")
RS_Daikon <- c("MYJO", "NELO", "TOBG", "WMBG")
RS_European <- c("DAJO", "ESNK", "SPEU", "CBBG")
RS_Oilseed <- c("AROL", "COOL", "ADOL")
RS_Rattail <- c("RABG", "RACA", "MABG")
Rostratus <- c("GMIL")
```

The paper has three main hypotheses about weed origin: The weeds could come from the crops, either directly or as hybrids; the weeds could be direct decendants of R.r. raphanistrum natives; or the weeds could be a hybrid of landra and R.r. raphanistrum natives.

#### Weeds from crops (In Paper)

If the weeds are feral crops, then they should be similar (not distinguishable by AMOVA).
This is comparing all crops as one group to all weeds as one group.



```{r AmovaWeedCrop_InPaper}

goodgroups <- c(RRR_Weed, RS_Daikon, RS_European, RS_Oilseed, RS_Rattail)
poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = goodgroups)

Allpopsamova <- poppr.amova(poppr_filter_ATA, ~Species/Pop, filter = T, 
                            missing = "ignore", within = FALSE)

Allpopsamova
amova.test <- randtest( Allpopsamova, nrepet = amovaReps) # Test for significance
plot(amova.test)
amova.test

```


#### Weeds from Native R.r. raphanistrum

If the weeds are most closely related to the native R.r. raphanistrum (this is the hypothesis we favor), then they shouldn't be significantly different from each other.


##### Weed - West

Here I'm comparing all weeds as one group, to native western R. r. raphanistrum as one group.


```{r AmovaWeedWest}

goodgroups <- c(RRR_Weed, RRR_Native_West)
poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = goodgroups)

Allpopsamova <- poppr.amova(poppr_filter_ATA, ~Group/Pop, filter = T, 
                            missing = "ignore", within = FALSE)
Allpopsamova
amova.test <- randtest( Allpopsamova, nrepet = amovaReps) # Test for significance
plot(amova.test)
amova.test

```

##### Weed - East

Here I'm comparing all weeds as one group, to native eastern R. r. raphanistrum as one group.

```{r AmovaWeedEast}

goodgroups <- c(RRR_Weed, RRR_Native_East)
poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = goodgroups)
Allpopsamova <- poppr.amova(poppr_filter_ATA, ~Group/Pop, filter = T, 
                            missing = "ignore", within = FALSE )
Allpopsamova
amova.test <- randtest( Allpopsamova, nrepet = amovaReps) # Test for significance
plot(amova.test)
amova.test

```


#### Weeds from hybridization of landra and native R.r. raphanistrum; In paper

If the weeds are hybrids of landra and native R.r. raphanistrum, then it should group with either in a paired analysis, however landra and native R.r. raphanistrum should be significantly different from one another.

Here I'm comparing all weeds as one group, to native western R. r. landra as one group.
Weed vs landra
(Weed vs native R.r. raphanistrum is above)

```{r AmovaWeedLandra_InPaper}

goodgroups <- c(RRR_Weed, Rr_Landra)
poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = goodgroups)
Allpopsamova <- poppr.amova(poppr_filter_ATA, ~Group/Pop, filter = T, 
                            missing = "ignore", within = FALSE )
Allpopsamova
amova.test <- randtest( Allpopsamova, nrepet = amovaReps) # Test for significance
plot(amova.test)
amova.test




```

##### Native R.r. raphanistrum vs landra

Here I'm comparing all R.r. landra as one group, to native western and eastern R. r. raphanistrum as one group.

```{r AmovaLandraRRR}

goodgroups <- c(RRR_Native_East, RRR_Native_West, Rr_Landra)
poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = goodgroups)

adegenet::strata(poppr_filter_ATA)$Group <- plyr::revalue(adegenet::strata(poppr_filter_ATA)$Group, c("raphNatW"="raphNatWE", "raphNatE"="raphNatWE"))

Allpopsamova <- poppr.amova(poppr_filter_ATA, ~Group/Pop, filter = T, 
                             missing = "ignore", within = FALSE)
Allpopsamova
amova.test <- randtest( Allpopsamova, nrepet = amovaReps) # Test for significance
plot(amova.test)
amova.test




```

##### Weed - East and West together; In paper

Here I'm comparing all weeds as one group, to native western and eastern R. r. raphanistrum, treated as a single group.

```{r AmovaWeedRRR_InPaper}

goodgroups <- c(RRR_Weed, RRR_Native_West, RRR_Native_East)

poppr_filter_ATA <- popsub(Allpops_filter_ATA, sublist = goodgroups)

Allpopsamova <- poppr.amova(poppr_filter_ATA, ~Group/Pop, filter = T, 
                            missing = "ignore", within = FALSE )
Allpopsamova
amova.test <- randtest( Allpopsamova, nrepet = amovaReps ) # Test for significance
plot(amova.test)
amova.test

```


### Using PCs instead of raw molecular data

```{r}
PCA.dat <- read.table("Marker.pca", skip=11)

PCA.ind <- read.table("../MungedData/Marker.ind")

colnames(PCA.ind) <- c("Individual", "U", "PopNum")

pca.lab <- cbind(PCA.ind, PCA.dat)

#Get other useful labels and join

labels.dat <- read.csv("../OriginalData/MarkerPopOrder.csv", header=F,
                       col.names=c("Individual", "Type", "Pop", "Order", "Name", "Species", "Color", "Vernalization", "DTF", "Bins", "locals"))

pca.lab <- full_join(pca.lab, labels.dat)

pca.lab$new.name <- factor( paste( pca.lab$Name, " (", pca.lab$Pop, ")", sep="" ) )
pca.lab <- droplevels(pca.lab)

```

#### ANOVA of just PC1, all groups

Like the first AMOVA above, I'm just asking whether the 9 groups are different from one another, but now I'm using PC1 as input. Since now this is a real linear model instead of whatever poppr does, I get both that they're different, and their deviations:

```{r}

mod <- lm( pca.lab$V1 ~ pca.lab$locals )
summary( mod )
car::Anova( mod )

```


#### MANOVA of PCs 1-10

PC1 doesn't capture all the differences between groups, so here I'm doing a MANOVA of PCs 1-10, which is ~30% of the variation in the molecular data (I just found that the numbers I was getting from SmartPCA weren't %variance explained like I thought, so I've recalculated it):

```{r}

pca.lab$locals <- relevel(pca.lab$locals, ref="raphNN")

Mmod <- lm( as.matrix( pca.lab[, 4:13] ) ~ pca.lab$locals )

car::Manova( Mmod )

```

The MANOVA shows that there is at least one significant difference between groups.

#### Long list of MANOVA model outputs

```{r}
summary(Mmod, multivariate=TRUE)

```

#### Mahalanobis

This table shows the coefficients from all ten linear models of the MANOVA. I've already added the intercept back in to each value, so they're easier to read.

```{r}
MmodCo <- Mmod$coefficients
MmodCo <- t(MmodCo)
MmodCo[,2:9] <- MmodCo[,2:9] + MmodCo[,1]
MmodCo <- data.frame(t(MmodCo))
row.names(MmodCo) <- c("NonNatives", "Daikon", "European", "lanmar", "Oilseed", "raphNatE", "raphNatW", "Rattail", "rostratus")
MmodCo

```

If I plot V1 vs V2, it looks essentially like a centroid-only version of the PCA. So I haven't broken anything.

```{r PCAgroupCentroid}

MmodCo$colors <- c(NonNatives, Daikon, European, lanmar, Oilseed, raphNatE, raphNatW, Rattail, rostratus)


plot(MmodCo$V1, MmodCo$V2, col=MmodCo$colors, pch=16, cex=1.2)

legend("bottomleft", legend = c("NonNatives", "Daikon", "European", "lanmar", "Oilseed", "raphNatE", "raphNatW", "Rattail", "rostratus"), col=MmodCo$colors, pch=16)

```


So, now you can see that, for instance, only rostratus and the weedy and native R.r. raphanistrum had negitive intercepts in PC1. Obviously, that's also what the original PCA plot says, but now, I have numbers.

I put this table into a Mahalanobis equation. It takes multi-dimensional data from related or unrelated measures, and converts it into directionless distances. So, it's taking the first 8 dimensions of orthoganal coordinates, and turning them into how far apart each group is from the rest. I can't use all 10 because the matrix is singular. I'm not sure if this is useful or not.

```{r}
#dim(MmodCo)
#
#qr(MmodCo)$rank
#
#
#mahalanobis.dist(MmodCo[,1:2])

S <-  var(MmodCo[,1:8])
mahalanobis(MmodCo[,1:8], center = F, cov = S)

```
