---
title: "PhenotypicAnalysis"
author: "Amanda Charbonneau"
date: "September 10, 2015"
output: 
  html_document: 
    fig_height: 10
    toc: yes
---

```{r SourcePackages, include=FALSE}
rm( list=ls())
source("~/Dropbox/RadishWeedEvolution/MarkerPaperMeta/Functionarium.R")
source("~/Dropbox/DworkinLabSharedMaterial/scripts/WRP_FUNCTIONS.R")
require(lme4) #modeling
#require(car) #ANOVA
require(RColorBrewer)
#require(MuMIn) #summary stats for LMER models
#require(arm) #extractAIC
require(MCMCglmm)
```


```{r ImportData, echo=FALSE}
FlowData <- read.csv("../output/CombinedDataSet.csv")
FlowData$FlowDate <- as.Date(FlowData$FlowDate)
FlowData$GermDate <- as.Date(FlowData$GermDate)
FlowData$PlantDate <- as.Date(FlowData$PlantDate)
FlowData$DOYPS <- ifelse(FlowData$DOYP < 78, FlowData$DOYP + 287, FlowData$DOYP - 78 )
FlowData$DOYGS <- ifelse(FlowData$DOYG < 78, FlowData$DOYG + 287, FlowData$DOYG - 78 )

```

Global parameters for MCMCglmm

```{r ModelParameters}

nitt <- 10000#100000   # Number of iterations
thin <- 50       # Thinning interval
burnin <- 5000   # burnin period
pr <- TRUE       # save posterior distribution of random effects

```

```{r RrrSubset, echo=FALSE}
# Data Subset for R.r. raphanistrum analysis
RrrData <- droplevels(FlowData[FlowData$Taxonomy == "raphanistrum",])
RrrData <- keepCols(RrrData, c("ID", "GTF", "Geography", "Seedstock", "GrowthEnvironment", "DOYGS", "DaysVernNoNA", "Pop", "Experiment", "Habitat", "PTF", "DOYPS", "Native", "PTG", "SurviveGTF"))

RrrData$ID <- row.names(RrrData)
RrrData$Geography <- relevel(RrrData$Geography, "west")
RrrData$Habitat <- relevel(RrrData$Habitat, "natural")
RrrData$Ag <- RrrData$Habitat
levels(RrrData$Ag) <- c("nonag", "ag", "nonag")
RrrData <- RrrData[!is.na(RrrData$SurviveGTF),] #Only keep data for individuals that have a flowering date
RrrData <- RrrData[!is.na( RrrData$DOYGS ),] #Only keep data for individuals that have a germination date
```

```{r RrNativeSubset, echo=FALSE}
# Data Subset for R.raphanistrum natives analysis
NatData <- droplevels(FlowData[FlowData$Geography != "nonNative" & FlowData$Species != "sativus",])
NatData <- keepCols(NatData, c("ID", "GTF", "Geography", "Seedstock", "GrowthEnvironment", "DOYGS", "DaysVernNoNA", "Pop", "Experiment", "Habitat", "PTF", "DOYPS", "Native", "Latency", "Species", "SubSpecies", "Taxonomy", "SurviveGTF"))

NatData$ID <- row.names(NatData)
NatData$Geography <- relevel(NatData$Geography, "west")
NatData$Habitat <- relevel(NatData$Habitat, "natural")
NatData$SubSpecies <- relevel(NatData$SubSpecies, "raphanistrum")
NatData <- NatData[!is.na(NatData$SurviveGTF),] #Only keep data for individuals that have a flowering date
NatData <- NatData[!is.na(NatData$DOYGS),]  #Only keep data for individuals that have a germination date
```
#Modeling

Only *R.r. raphanistrum*
Response variable: Days from germination to flowering
Here, I'm using germination to flowering (or last alive date) as the response, which will be more accurate than planting to flowering, but have many more NA's.

```{r RrrModels, echo=TRUE, results="hide"}
R
Rrr_lmer <- lmer( SurviveGTF ~ Geography + Seedstock + DOYGS + DaysVernNoNA + (1|Experiment) + (1|Geography:Pop), data=RrrData  )

Rrr_glmm <- MCMCglmm( fixed = SurviveGTF ~ Geography  + Seedstock + DOYGS + DaysVernNoNA, 
                      random = ~ Experiment + Geography:Pop,
                      nitt = nitt,
                      thin = thin,
                      burnin = burnin,
                      pr = pr, #save posterior distribution of random effects
                      data = RrrData)

```


Only native range *R.raphanistrum*
#This is the model that is roughly equivalent to the model we used to look at the differences between the ***R.r. raphanistrum*** plants, just using **SubSpecies** instead of **Geography**, and using the **SurviveGTF** response. 

```{r RrNatModels, echo=TRUE, results="hide"}

Nat_lmer <- lmer( SurviveGTF ~ SubSpecies + Seedstock + DaysVernNoNA + (1|Experiment) + (1|SubSpecies:Pop), data=NatData  )


Nat_glmm <- MCMCglmm(fixed = SurviveGTF ~ SubSpecies + Seedstock + DaysVernNoNA, 
                     random = ~ Experiment + SubSpecies:Pop, 
                     nitt = nitt,
                     thin = thin,
                     burnin = burnin,
                     pr = pr, #save posterior distribution of random effects
                     data = NatData)
```

##Model output for just *R.r. raphanistrum*
```{r ModelStatsRrr, echo=FALSE}

summary( Rrr_glmm )
plot( Rrr_glmm$Sol )
acf( Rrr_glmm$VCV )

```

##Model output for just native range *R. raphanistrum*
```{r ModelStatsNat, echo=FALSE}

summary( Nat_glmm )
plot( Nat_glmm$Sol )
acf( Nat_glmm$VCV )

```

pop_chains <- Rrr_glmm$Sol[, 14:29] + Rrr_glmm$Sol[,1]
colMeans(pop_chains)
HPDinterval(pop_chains)




